---
title: "1_1_Seurat object setup"
output:
  pdf_document:
header-includes:
     - \pagenumbering{gobble}
     - \usepackage{etoolbox}
     - \preto{\section}{\ifhmode\unskip\fi}
     - \usepackage{titling}
     - \pretitle{\begin{center}\LARGE\color{blue}\textbf}
     - \posttitle{\par\end{center}\vskip 0.5em}
     - \usepackage{sectsty}
     - \sectionfont{\color{orange}\fontsize{18}{20}\selectfont}
     - \subsectionfont{\color{purple}\fontsize{14}{16}\selectfont}
date: "`r Sys.Date()`"
---

```{r message=FALSE, warning=FALSE, include=FALSE}
# R version 4.2.2
if(!require(Matrix))remotes::install_version("Matrix", version = "1.5-1")
if(!require(SeuratObject))remotes::install_version("SeuratObject", version = "4.1.3")
if(!require(Seurat))remotes::install_version("Seurat", version = "4.3.0")
if(!require(clustree))BiocManager::install("clustree")
if(!require(patchwork))install.packages("patchwork")
if(!require(knitr))install.packages("knitr")
if(!require(dplyr))install.packages("dplyr")
if(!require(ggplot2))BiocManager::install("ggplot2")
if(!require(writexl))install.packages("writexl")
if(!require(readxl))install.packages("readxl")
#for first time run# install.packages("tinytex")
#for first time run# tinytex::install_tinytex()
```

```{r message=FALSE, warning=FALSE, include=FALSE}
four.good.colors <- c("#E84D8A", "#FEB326", "#64C5EB", "#7F58AF")
eight.good.colors <- c("#1F77B4","#FF7F0E","#2CA02C","#D62728","#9467BD","#8C564B","#BCBD22","#17BECF")
ten.good.colors <- c("#1F77B4","#FF7F0E","#2CA02C","#D62728","#9467BD",
                     "#8C564B","#E377C2","#7F7F7F","#BCBD22","#17BECF")
eleven.good.colors <- c("#1F77B4","#FF7F0E","#2CA02C","#D62728","#9467BD",
                     "#8C564B","#E377C2","#BCBD22","#17BECF","#440154","#FDE725")
heatmap.five.colors <- c("#440154","#3B528B","#21908C","#5DC863","#FDE725")
```

# Sample9-16 filter threshold setting: 
## Include cells where at least this many features(i.e. genes) are detected: 50
## Include features(genes) detected in at least this many cells: 6
## nFeature_RNA = gene number < 11000 & > 3000
## nCount_RNA = RNA number < 150000
### High nFeature_RNA might mean doublet cells; low nFeature_RNA might mean thie cell might be bad or not seq enough.
## percent.mt = the ratio of mitochondria filtered 10%
### High ratio of mitochondria RNA means the cell is not active enough.
\pagebreak
```{r echo=FALSE, fig.height=3.8, fig.width=6, message=FALSE, warning=FALSE, results='asis'}
sample.id <- 9:16
sample.name <- c(rep('black', (sample.id[1]-1)), paste0("Sample", 9:12, ".WT"), paste0("Sample", 13:16, ".Mut"))
colors <- c(rep('black', (sample.id[1]-1)), eight.good.colors)
nCount <- c()
nFeature <- c()
percent <- c()
Replicate <- c()

for (i in sample.id) {
X10 <- Read10X(data.dir = paste0("./GEX_Kristen_", 
  i, "_Di2023/outs/filtered_feature_bc_matrix"))
raw <- CreateSeuratObject(counts = X10, 
    project = sample.name[i], min.features = 50, min.cells = 6)

raw[["percent.mt"]] <- PercentageFeatureSet(raw, pattern = "^mt-")

VlnPlot <- VlnPlot(raw, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol =3, cols = colors[i]) + 
  labs(subtitle = '←Before filtering') + theme(plot.subtitle = element_text(size = 6))
print(VlnPlot)

FeatureScatter1 <- FeatureScatter(raw, feature1 = "nCount_RNA", feature2 = "nFeature_RNA", 
               cols = colors[i]) + theme(legend.position = "none") + ggtitle(paste0(sample.name[i], '-before'))

raw <- subset(raw, subset = nFeature_RNA > 3000 & nFeature_RNA < 11000 & nCount_RNA < 150000 & percent.mt < 10)

VlnPlot2 <- VlnPlot(raw, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol =3, 
                    cols = colors[i]) + labs(subtitle = '←After filtering') + 
  theme(plot.subtitle = element_text(size = 6))
print(VlnPlot2)

cat("  \n")
cat(paste0("The cell numbers before filterring of", sample.name[i], " : ", ncol(X10)))
cat("  \n")

cat(paste0("The final cell number after filtering of", sample.name[i], " : ", ncol(as.data.frame(raw[["RNA"]]@counts))))    
cat("  \n")

cat("\n\n\\pagebreak\n")
FeatureScatter2 <- FeatureScatter(raw, feature1 = "nCount_RNA", feature2 = "nFeature_RNA", 
               cols = colors[i]) + theme(legend.position = "none") + ggtitle(paste0(sample.name[i], '-after'))
print(FeatureScatter1 + FeatureScatter2)

nCount_RNA <- raw@meta.data$nCount_RNA
nFeature_RNA <- raw@meta.data$nFeature_RNA
percent.mt <- raw@meta.data$percent.mt
nCount <- c(nCount, nCount_RNA)
nFeature <- c(nFeature, nFeature_RNA)
percent <- c(percent, percent.mt)
Replicate <- c(Replicate, rep(sample.name[i], length(nCount_RNA)))

assign(paste0(sample.name[i], '.raw'), raw)
cat("\n\n\\pagebreak\n")}
```

\pagebreak
# Reproducibility of nCount, nFeature, percent.mt
```{r echo=FALSE, fig.height=4.5, fig.width=6, message=FALSE, warning=FALSE, results='asis'}
Replicate <- factor(Replicate, levels = c(paste0("Sample", 9:12, ".WT"), paste0("Sample", 13:16, ".Mut")), ordered = TRUE)
meta.data.merge <- data.frame(nCount, nFeature, percent, Replicate)

ggplot(meta.data.merge,aes(Replicate, nCount))+
  ggdist::stat_halfeye(mapping = aes(fill=Replicate),width = 0.6, .width = 0, 
  justification = -0.4, point_colour = NA,alpha=1) + 
  geom_jitter(mapping = aes(color=Replicate),width = 0.15, alpha = 0.4,size=0.05)+
  geom_boxplot(width = 0.15, outlier.shape = NA,fill=NA,size=0.5)+
  labs(title = "nCount_RNA")+scale_y_continuous(expand = c(0.02,0))+
  scale_x_discrete(expand = c(0,0.3))+scale_fill_manual(values = colors[9:16])+
  scale_color_manual(values = colors[9:16]) +theme_classic()+
  theme(axis.text.x.bottom = element_text(angle = 45,hjust = 1,size = 10,color = "black"),
    axis.text.y.left = element_text(size = 12,color = "black"),axis.title.y.left = element_blank(),
    axis.title.x.bottom = element_blank(),axis.ticks.length=unit(0.2, "cm"),
    plot.title = element_text(size = 18,hjust = 0.5),legend.position = "none")
cat("\n\n\\pagebreak\n")

ggplot(meta.data.merge,aes(Replicate, nFeature))+
  ggdist::stat_halfeye(mapping = aes(fill=Replicate),width = 0.6, .width = 0, 
  justification = -0.4, point_colour = NA,alpha=1) + 
  geom_jitter(mapping = aes(color=Replicate),width = 0.15, alpha = 0.4,size=0.05)+
  geom_boxplot(width = 0.15, outlier.shape = NA,fill=NA,size=0.5)+
  labs(title = "nFeature_RNA")+scale_y_continuous(expand = c(0.02,0))+
  scale_x_discrete(expand = c(0,0.3))+scale_fill_manual(values = colors[9:16])+
  scale_color_manual(values = colors[9:16]) +theme_classic()+
  theme(axis.text.x.bottom = element_text(angle = 45,hjust = 1,size = 10,color = "black"),
    axis.text.y.left = element_text(size = 12,color = "black"),axis.title.y.left = element_blank(),
    axis.title.x.bottom = element_blank(),axis.ticks.length=unit(0.2, "cm"),
    plot.title = element_text(size = 18,hjust = 0.5),legend.position = "none")
cat("\n\n\\pagebreak\n")

ggplot(meta.data.merge,aes(Replicate, percent))+
  ggdist::stat_halfeye(mapping = aes(fill=Replicate),width = 0.6, .width = 0, 
  justification = -0.4, point_colour = NA,alpha=1) + 
  geom_jitter(mapping = aes(color=Replicate),width = 0.15, alpha = 0.4,size=0.05)+
  geom_boxplot(width = 0.15, outlier.shape = NA,fill=NA,size=0.5)+
  labs(title = "nFeature_RNA")+scale_y_continuous(expand = c(0.02,0))+
  scale_x_discrete(expand = c(0,0.3))+scale_fill_manual(values = colors[9:16])+
  scale_color_manual(values = colors[9:16]) +theme_classic()+
  theme(axis.text.x.bottom = element_text(angle = 45,hjust = 1,size = 10,color = "black"),
    axis.text.y.left = element_text(size = 12,color = "black"),axis.title.y.left = element_blank(),
    axis.title.x.bottom = element_blank(),axis.ticks.length=unit(0.2, "cm"),
    plot.title = element_text(size = 18,hjust = 0.5),legend.position = "none")
```

\pagebreak
```{r message=FALSE, warning=FALSE, include=FALSE}
for (i in sample.id) {
name <- paste0(sample.name[i], '.raw')
raw <- get(name)
raw <- NormalizeData(raw, normalization.method = "LogNormalize", scale.factor = 10000)
#normalizes the feature expression measurements for each cell by the total expression
raw <- FindVariableFeatures(raw, selection.method = "vst", nfeatures = 2000)  #2000 HVGs
assign(paste0(sample.name[i], '.raw'), raw)
}
```

```{r message=FALSE, warning=FALSE, include=FALSE}
WT.anchors <- FindIntegrationAnchors(object.list = list(Sample9.WT.raw, 
              Sample10.WT.raw,Sample11.WT.raw,Sample12.WT.raw), dims = 1:20)
WT.integrated <- IntegrateData(anchorset = WT.anchors, dims = 1:20)
saveRDS(WT.integrated,"./WT.integrated.22.03.2024.rds")
# treat all WT as replicates.

merge.anchors <- FindIntegrationAnchors(object.list = list(Sample9.WT.raw, 
              Sample10.WT.raw,Sample11.WT.raw,Sample12.WT.raw,Sample13.Mut.raw,
              Sample14.Mut.raw,Sample15.Mut.raw,Sample16.Mut.raw), dims = 1:20)
merge.integrated <- IntegrateData(anchorset = merge.anchors, dims = 1:20)
saveRDS(merge.integrated,"./merge.integrated.22.03.2024.rds")
# deeply normalize all the eight samples
```

# Merge all and treat all eight WT / Mut as replicates
## did deeply normalize WT and Mut, as well as seq depth correct
```{r message=FALSE, warning=FALSE, include=FALSE}
# merge.integrated <- readRDS("./merge.integrated.26.06.2023.rds")
# WT.integrated <- readRDS("./WT.integrated.22.03.2024.rds")

# Find 2000 highly variable genes (HVGs)
ALL.merge.raw <- FindVariableFeatures(merge.integrated, selection.method = "vst", nfeatures = 2000)  #2000 HVGs
WT.merge.raw <- FindVariableFeatures(WT.integrated, selection.method = "vst", nfeatures = 2000)  #2000 HVGs

## gene expression normalization
ALL.merge.raw <- ScaleData(ALL.merge.raw, features = rownames(ALL.merge.raw))
WT.merge.raw <- ScaleData(WT.merge.raw, features = rownames(WT.merge.raw))
```

```{r echo=FALSE, fig.height=3.8, fig.width=6, message=FALSE, warning=FALSE}
# Show the PCA results of the 2000 HVGs
## PC1 as an example
ALL.merge.raw <- RunPCA(ALL.merge.raw, features = rownames(ALL.merge.raw))    
WT.merge.raw <- RunPCA(WT.merge.raw, features = rownames(WT.merge.raw))   
# print(ALL.merge.raw[["pca"]], dims = 1:5, nfeatures = 5)    #PCA results 
# VizDimLoadings(ALL.merge.raw, dims = 1:2, reduction = "pca") 

Idents(ALL.merge.raw) <- factor(Idents(ALL.merge.raw), levels = levels(Replicate))   # reorder Seurat Idents levels.
Idents(WT.merge.raw) <- factor(Idents(WT.merge.raw), levels = levels(Replicate)[1:4])

DimPlot(ALL.merge.raw, reduction = "pca", cols = eight.good.colors, 
     label.size = 0.3) + theme(plot.title = element_text(size = 12)) + 
    theme(legend.position = "bottom", legend.text = element_text(size = 5)) + 
  labs(title = "Normalisation all 8 samples together", subtitle = 'deeply normalize WT & Mut + depth correct')
DimPlot(WT.merge.raw, reduction = "pca", cols = eight.good.colors[1:4], 
     label.size = 0.3) + theme(plot.title = element_text(size = 12)) + 
    theme(legend.position = "bottom", legend.text = element_text(size = 5)) + 
  labs(title = "Normalisation all 4 WT samples", subtitle = 'deeply normalize WT & Mut + depth correct')
# DimHeatmap(ALL.merge.raw, dims = 1:15, cells = 500, balanced = TRUE) 
```

```{r fig.height=3.5, fig.width=5, message=FALSE, warning=FALSE, include=FALSE, results='asis'}
# Decide the suitable PC we would use 
## This would help decide how much PC we need. Here we selected the top 12 PC. 
ALL.merge.raw <- JackStraw(ALL.merge.raw, num.replicate = 100,  dims = 20) 
ALL.merge <- ScoreJackStraw(ALL.merge.raw, dims = 1:20) 
# JackStrawPlot(ALL.merge, dims = 1:20) 

WT.merge.raw <- JackStraw(WT.merge.raw, num.replicate = 100,  dims = 20) 
WT.merge <- ScoreJackStraw(WT.merge.raw, dims = 1:20) 
# JackStrawPlot(ALL.merge, dims = 1:20)
```

```{r echo=FALSE, fig.height=3.5, fig.width=5, message=FALSE, warning=FALSE, results='asis'}
ElbowPlot(ALL.merge) + labs(title = "Normalisation all 8 samples together")
cat("  \n")
ElbowPlot(WT.merge) + labs(title = "Normalisation 4 WT samples")
```

\pagebreak
# Deciding the number of clusters for all eight merge
## decided the dim region 20, which is the most stable
### in dim 20, select resolution 0.34
```{r message=FALSE, warning=FALSE, include=FALSE}
para = c(0.02)
for (x in 2:18) {
  y = 0.02 * x
  para <- c(para, y)
}
ALL.merge <- ScoreJackStraw(ALL.merge.raw, dims = 1:20) 
ALL.merge <- FindNeighbors(ALL.merge, dims = 1:20)  
ALL.merge <- FindClusters(ALL.merge, resolution = para)
```
```{r echo=FALSE, fig.height=8, fig.width=7, message=FALSE, warning=FALSE}
clustree(ALL.merge)  #decided the dim region 20, which is the most stable. in 20, select resolution 0.34
```

\pagebreak
# Deciding the number of clusters for WT four merge
## decided the dim region 18, which is the most stable
### in dim 18, select resolution 0.18
```{r message=FALSE, warning=FALSE, include=FALSE}
para = c(0.02)
for (x in 2:14) {
  y = 0.02 * x
  para <- c(para, y)
}

WT.merge <- FindNeighbors(WT.merge, dims = 1:18)  
WT.merge <- FindClusters(WT.merge, resolution = para)
```
```{r echo=FALSE, fig.height=8, fig.width=7, message=FALSE, warning=FALSE}
clustree(WT.merge)  #decided the dim region 18, which is the most stable. in 18, select resolution 0.18
```

```{r message=FALSE, warning=FALSE, include=FALSE}
ALL.merge <- FindClusters(ALL.merge, resolution = 0.34)  ###basing on the above clustree result
ALL.merge <- RunTSNE(ALL.merge, dims = 1:20) 
ALL.merge <- RunUMAP(ALL.merge, dims = 1:20)

WT.merge <- FindClusters(WT.merge, resolution = 0.18)  ###basing on the above clustree result
WT.merge <- RunTSNE(WT.merge, dims = 1:18) 
WT.merge <- RunUMAP(WT.merge, dims = 1:18)
```

```{r message=FALSE, warning=FALSE, include=FALSE}
saveRDS(ALL.merge,"./ALL.merge.22.03.2024.rds")
saveRDS(WT.merge,"./R analysis/WT.merge.22.03.2024.rds")
```

# Nine clusters in four merged WT samples
## with unsplit cluster DE/Node/APS
```{r echo=FALSE, fig.height=7, fig.width=8, message=FALSE, warning=FALSE, results='asis'}
# WT.merge <- readRDS("./WT.merge.26.06.2023.rds")
# ALL.merge <- readRDS("./ALL.merge.26.06.2023.rds")
WT.merge <- RenameIdents(object = WT.merge, '0' = 'NasMeso', '1' = 'Epi', '2' = 'PS',
         '3' = 'ExVE', '4' = 'ExE Meso', '5' = 'EmVE', '6' = 'DE/Node/APS', '7' = 'ExE Ectoderm',
         '8' = 'PGCs')

DimPlot(WT.merge, reduction = "umap", label = T,label.color = "black",repel = T,label.size = 3.5,
        cols = eleven.good.colors) + 
  labs(title = "WT - clusters", subtitle = 'merge for WT samples')
```
\pagebreak
```{r fig.height=7, fig.width=8, message=FALSE, warning=FALSE, include=FALSE}
DE.Node.APS.raw <- subset(x = WT.merge, idents = c("DE/Node/APS"), invert=F)
DE.Node.APS.raw <- JackStraw(DE.Node.APS.raw, num.replicate = 100) 
DE.Node.APS <- ScoreJackStraw(DE.Node.APS.raw, dims = 1:20) 
# ElbowPlot(DE.Node.APS)

para = c(0.02)
for (x in 2:14) {
  y = 0.02 * x
  para <- c(para, y)
}
DE.Node.APS <- FindNeighbors(DE.Node.APS, dims = 1:12)  
DE.Node.APS <- FindClusters(DE.Node.APS, resolution = para)
# clustree(DE.Node.APS)  
DE.Node.APS <- FindClusters(DE.Node.APS, resolution = 0.16) 
DE.Node.APS <- RenameIdents(object = DE.Node.APS, '0' = 'DE', '1' = 'Node', '2' = 'APS')
DE.Node.APS <- RunTSNE(DE.Node.APS, dims = 1:12) 
DE.Node.APS <- RunUMAP(DE.Node.APS, dims = 1:12)
```

# Three split clusters in four merged WT samples
## in the cluster DE/Node/APS
```{r echo=FALSE, fig.height=7, fig.width=8, message=FALSE, warning=FALSE, results='asis'}
DimPlot(DE.Node.APS, reduction = "umap", label = T,label.color = "black",repel = T,label.size = 5,
        cols = c(eleven.good.colors[which(levels(Idents(WT.merge)) == "DE/Node/APS")],
        eleven.good.colors[length(eleven.good.colors)-1], eleven.good.colors[length(eleven.good.colors)])) + 
  labs(title = "WT - clusters", subtitle = 'DE.Node.APS split clusters')
```

```{r fig.height=7, fig.width=8, message=FALSE, warning=FALSE, include=FALSE}
matches <- grep("TRUE", levels(Idents(WT.merge)) != "DE/Node/APS")
other.clusters.name <- c()
for (i in matches) {
  clusters.name <- levels(Idents(WT.merge))[i]
  other.clusters.name <- c(other.clusters.name, clusters.name)
}
other.clusters <- subset(x = WT.merge, idents = other.clusters.name, invert=F)

WT.merge.anno <- merge(other.clusters, DE.Node.APS)
WT.merge.anno@graphs <- WT.merge@graphs
WT.merge.anno@reductions <- WT.merge@reductions
WT.merge.anno@commands <- WT.merge@commands
WT.merge.anno@tools <- WT.merge@tools
WT.merge.anno@meta.data <- WT.merge@meta.data
WT.merge.anno@assays <- WT.merge@assays
```

# Eleven clusters in four merged WT samples
## with split cluster DE/Node/APS
```{r echo=FALSE, fig.height=7, fig.width=8, message=FALSE, warning=FALSE, results='asis'}
WT.colors <- eleven.good.colors 

names(WT.colors) <- c(levels(Idents(WT.merge))[1:(which(levels(Idents(WT.merge)) == "DE/Node/APS")-1)], levels(Idents(DE.Node.APS))[1], levels(Idents(WT.merge))[(which(levels(Idents(WT.merge)) == "DE/Node/APS")+1):length(levels(Idents(WT.merge)))], levels(Idents(DE.Node.APS))[2:3])

DimPlot(WT.merge.anno, reduction = "umap", label = T,label.color = "black",repel = T,label.size = 3.5,
        cols = WT.colors) + 
  labs(title = "WT - clusters", subtitle = 'in four merged WT samples')

saveRDS(WT.merge.anno,"./WT.merge.anno.22.03.2024.rds")
```

\pagebreak
# Eleven clusters in all 8 samples merged together
## WT + Mut
```{r echo=FALSE, fig.height=7, fig.width=8, message=FALSE, warning=FALSE, results='asis'}
ALL.merge <- RenameIdents(object = ALL.merge, '0' = 'Epi', '1' = 'NasMes', '2' = 'PS', '3' = 'ExVE', '4' = 'ExE Mes', '5' = 'DE', '6' = 'Amnion/Chorion', '7' = 'ExE Ecto', '8' = 'EmVE', '9' = "Blood prog", "10" = "Node")

ALL.colors <- eleven.good.colors 
names(ALL.colors) <- levels(Idents(ALL.merge))

DimPlot(ALL.merge, reduction = "umap", label = T,label.color = "black",repel = T,label.size = 3.5,
        cols = ALL.colors) + 
  labs(title = "All - clusters", subtitle = 'merge all WT + Mut eight samples')
```

\pagebreak
## with split DE & Node clusters from 2 to 3
## with split ExE Mes cluster from 1 to 2
```{r fig.height=7, fig.width=8, message=FALSE, warning=FALSE, include=FALSE}
DE.Node.raw <- subset(x = ALL.merge, idents = c("DE", 'Node'), invert=F)
DE.Node.raw <- JackStraw(DE.Node.raw, num.replicate = 100) 
DE.Node <- ScoreJackStraw(DE.Node.raw, dims = 1:20) 
# ElbowPlot(DE.Node.APS)

para = c(0.02)
for (x in 2:14) {
  y = 0.02 * x
  para <- c(para, y)
}
DE.Node <- FindNeighbors(DE.Node, dims = 1:10)  
DE.Node <- FindClusters(DE.Node, resolution = para)
# clustree(DE.Node)  
DE.Node <- FindClusters(DE.Node, resolution = 0.14) 
DE.Node <- RenameIdents(object = DE.Node, '0' = 'Node', '1' = 'DE', '2' = 'APS')
DE.Node <- RunTSNE(DE.Node, dims = 1:10) 
DE.Node <- RunUMAP(DE.Node, dims = 1:10)

### ExE Mes

ExE.Mes.raw <- subset(x = ALL.merge, idents = c("ExE Mes"), invert=F)
ExE.Mes.raw <- JackStraw(ExE.Mes.raw, num.replicate = 100) 
ExE.Mes <- ScoreJackStraw(ExE.Mes.raw, dims = 1:20) 
# ElbowPlot(ExE.Mes.APS)

para = c(0.02)
for (x in 2:14) {
  y = 0.02 * x
  para <- c(para, y)
}
ExE.Mes <- FindNeighbors(ExE.Mes, dims = 1:10)  
ExE.Mes <- FindClusters(ExE.Mes, resolution = para)
# clustree(ExE.Mes)  
ExE.Mes <- FindClusters(ExE.Mes, resolution = 0.08) 
ExE.Mes <- RenameIdents(object = ExE.Mes, '0' = 'ExE Mes', '1' = 'PGCs')
ExE.Mes <- RunTSNE(ExE.Mes, dims = 1:10) 
ExE.Mes <- RunUMAP(ExE.Mes, dims = 1:10)
```

```{r eval=FALSE, fig.height=7, fig.width=8, message=FALSE, warning=FALSE, include=FALSE, results='asis'}
DimPlot(DE.Node, reduction = "umap", label = T,label.color = "black",repel = T,label.size = 5) + 
  labs(title = "All - clusters", subtitle = 'DE.Node split clusters')
cat("  \n")
DimPlot(ExE.Mes, reduction = "umap", label = T,label.color = "black",repel = T,label.size = 5) + 
  labs(title = "All - clusters", subtitle = 'ExE Mes split clusters')
```

```{r fig.height=7, fig.width=8, message=FALSE, warning=FALSE, include=FALSE}
matches <- intersect(intersect(grep("TRUE", levels(Idents(ALL.merge)) != "DE"), 
            grep("TRUE", levels(Idents(ALL.merge)) != "Node")), grep("TRUE", levels(Idents(ALL.merge)) != "ExE Mes"))

other.clusters.name <- c()
for (i in matches) {
  clusters.name <- levels(Idents(ALL.merge))[i]
  other.clusters.name <- c(other.clusters.name, clusters.name)
}
other.clusters <- subset(x = ALL.merge, idents = other.clusters.name, invert=F)

ALL.merge.anno <- merge(other.clusters, c(ExE.Mes, DE.Node))
ALL.merge.anno@graphs <- ALL.merge@graphs
ALL.merge.anno@reductions <- ALL.merge@reductions
ALL.merge.anno@commands <- ALL.merge@commands
ALL.merge.anno@tools <- ALL.merge@tools
ALL.merge.anno@meta.data <- ALL.merge@meta.data
ALL.merge.anno@assays <- ALL.merge@assays
```

# 13 clusters in All 8 merged samples
## with split cluster DE/Node/APS and EXE.Mes/PGC
```{r echo=FALSE, fig.height=7, fig.width=8, message=FALSE, warning=FALSE, results='asis'}
extra.colors <- c("#006400","#7F7F7F")
names(extra.colors) <- c('PGCs','APS' )
ALL.colors <- c(ALL.colors, extra.colors)
saveRDS(ALL.colors,"./ALL.colors.22.03.2024.rds")

ALL.merge.anno$seurat_clusters <- Idents(ALL.merge.anno)

DimPlot(ALL.merge.anno, reduction = "umap", label = T,label.color = "black",repel = T,label.size = 3.5, cols = ALL.colors) + 
  labs(title = "All - clusters", subtitle = 'Final merge all WT + Mut eight samples')
cat("\n\n\\pagebreak\n")

saveRDS(ALL.merge.anno,"./ALL.merge.anno.22.03.2024.rds")
# ALL.merge.anno <- readRDS("./ALL.merge.anno.22.03.2024.rds")
```

```{r echo=FALSE, fig.height=7, fig.width=8, message=FALSE, warning=FALSE, results='asis'}
cat('# Cell number in each cluster')
merge <- ALL.merge.anno

cell.counts <- table(merge$seurat_clusters)
cell.count.df <- data.frame(Cluster = names(cell.counts), 'Cell.Count' = as.numeric(cell.counts))
kable(cell.count.df)
cat("\n\n\\pagebreak\n")
```

# Biological replicates?
```{r echo=FALSE, fig.height=6, fig.width=7, message=FALSE, warning=FALSE, results='asis'}
Idents(ALL.merge.anno) <- ALL.merge.anno$orig.ident
DimPlot(ALL.merge.anno, reduction = "umap", cols = eight.good.colors) + 
  labs(title = "All eight WT + Mut samples") + ggplot2::theme(legend.position="top")
cat("\n\n\\pagebreak\n")
DimPlot(ALL.merge.anno, reduction = "umap", cols = c(rep(eight.good.colors[1],4), rep(eight.good.colors[2],4))) + 
  labs(title = "Compare WT and Mut samples") + ggplot2::theme(legend.position="top")
```

\pagebreak
```{r echo=FALSE, fig.height=6.5, fig.width=6.5, message=FALSE, warning=FALSE, results='asis'}
for (i in sample.id) {
paste0(sample.name[i])
subset <- subset(x = ALL.merge.anno, idents = c(sample.name[i]), invert=F)
assign(paste0(sample.name[i]), subset)}

DimPlot(Sample9.WT, reduction = "umap", cols = eight.good.colors[1], pt.size = 0.0001) + ggplot2::theme(legend.position="top") +
 (DimPlot(Sample10.WT, reduction = "umap", cols = eight.good.colors[2], pt.size = 0.0001) + ggplot2::theme(legend.position="top")) +
 (DimPlot(Sample11.WT, reduction = "umap", cols = eight.good.colors[3], pt.size = 0.0001) + ggplot2::theme(legend.position="top")) +
 (DimPlot(Sample12.WT, reduction = "umap", cols = eight.good.colors[4], pt.size = 0.0001) + ggplot2::theme(legend.position="top")) +
 (DimPlot(Sample13.Mut, reduction = "umap", cols = eight.good.colors[5], pt.size = 0.0001) + ggplot2::theme(legend.position="top")) +
 (DimPlot(Sample14.Mut, reduction = "umap", cols = eight.good.colors[6], pt.size = 0.0001) + ggplot2::theme(legend.position="top")) +
 (DimPlot(Sample15.Mut, reduction = "umap", cols = eight.good.colors[7], pt.size = 0.0001) + ggplot2::theme(legend.position="top")) +
 (DimPlot(Sample16.Mut, reduction = "umap", cols = eight.good.colors[8], pt.size = 0.0001) + ggplot2::theme(legend.position="top"))
cat("\n\n\\pagebreak\n")

nFeature_numbers <- c(Sample9.WT@meta.data[["nFeature_RNA"]],Sample10.WT@meta.data[["nFeature_RNA"]],
                    Sample11.WT@meta.data[["nFeature_RNA"]],Sample12.WT@meta.data[["nFeature_RNA"]],
                    Sample13.Mut@meta.data[["nFeature_RNA"]],Sample14.Mut@meta.data[["nFeature_RNA"]],
                    Sample15.Mut@meta.data[["nFeature_RNA"]],Sample16.Mut@meta.data[["nFeature_RNA"]])
numeric_vector <- as.numeric(nFeature_numbers)
average_of_nFeature_RNA <- round(mean(numeric_vector))

nCount_numbers <- c(Sample9.WT@meta.data[["nCount_RNA"]],Sample10.WT@meta.data[["nCount_RNA"]],
                      Sample11.WT@meta.data[["nCount_RNA"]],Sample12.WT@meta.data[["nCount_RNA"]],
                      Sample13.Mut@meta.data[["Count_RNA"]],Sample14.Mut@meta.data[["nCount_RNA"]],
                      Sample15.Mut@meta.data[["Count_RNA"]],Sample16.Mut@meta.data[["nCount_RNA"]])
numeric_vector2 <- as.numeric(nCount_numbers)
average_of_nCount_RNA <- round(mean(numeric_vector2))
```

# 13 clusters
```{r echo=FALSE, fig.height=7, fig.width=8, message=FALSE, warning=FALSE, results='asis'}
Idents(ALL.merge.anno) <- ALL.merge.anno$orig.ident
WT <- subset(x = ALL.merge.anno, idents = sample.name[9:12], invert=F)
Mut <- subset(x = ALL.merge.anno, idents = sample.name[13:16], invert=F)
Mut13 <- subset(x = ALL.merge.anno, idents = sample.name[13], invert=F)
Mut14 <- subset(x = ALL.merge.anno, idents = sample.name[14], invert=F)
Mut15 <- subset(x = ALL.merge.anno, idents = sample.name[15], invert=F)
Mut16 <- subset(x = ALL.merge.anno, idents = sample.name[16], invert=F)

Idents(ALL.merge.anno) <- ALL.merge.anno$seurat_clusters
Idents(WT) <- WT$seurat_clusters
Idents(Mut) <- Mut$seurat_clusters
Idents(Mut13) <- Mut13$seurat_clusters
Idents(Mut14) <- Mut14$seurat_clusters
Idents(Mut15) <- Mut15$seurat_clusters
Idents(Mut16) <- Mut16$seurat_clusters

DimPlot(WT, reduction = "umap", label = T,label.color = "black",repel = T,label.size = 4,
        cols = ALL.colors) + 
  labs(title = "Clusters in WT merged group", subtitle = 'merge all WT + Mut eight samples')

cell.number.WT <- c()
cat("  \n")
for (i in levels(Idents(WT))){
cat(paste0('The cell number of cluster', i, ' in WT group is ', sum(Idents(WT) == i)), "\n")
cat("  \n")
cell.number.WT <- c(cell.number.WT, sum(Idents(WT) == i))}

cat("\n\n\\pagebreak\n")
DimPlot(Mut, reduction = "umap", label = T,label.color = "black",repel = T,label.size = 4,
        cols = ALL.colors) + 
  labs(title = "Clusters in Mut merged group", subtitle = 'merge all WT + Mut eight samples')

cell.number.Mut <- c()
cat("  \n")
for (i in levels(Idents(Mut))){
cat(paste0('The cell number of cluster', i, ' in Mut group is ', sum(Idents(Mut) == i)), "\n")
cat("  \n")
cell.number.Mut <- c(cell.number.Mut, sum(Idents(Mut) == i))}

cat("\n\n\\pagebreak\n")

DimPlot(Mut13, reduction = "umap", label = T,label.color = "black",repel = T,label.size = 4,
        cols = ALL.colors) + 
  labs(title = "Clusters in Mut13", subtitle = 'merge all WT + Mut eight samples')

cell.number.Mut13 <- c()
cat("  \n")
for (i in levels(Idents(Mut13))){
cat(paste0('The cell number of cluster', i, ' in Mut13 is ', sum(Idents(Mut13) == i)), "\n")
cat("  \n")
cell.number.Mut13 <- c(cell.number.Mut13, sum(Idents(Mut13) == i))}

cat("\n\n\\pagebreak\n")
DimPlot(Mut14, reduction = "umap", label = T,label.color = "black",repel = T,label.size = 4,
        cols = ALL.colors) + 
  labs(title = "Clusters in Mut14", subtitle = 'merge all WT + Mut eight samples')

cell.number.Mut14 <- c()
cat("  \n")
for (i in levels(Idents(Mut14))){
cat(paste0('The cell number of cluster', i, ' in Mut14 is ', sum(Idents(Mut14) == i)), "\n")
cat("  \n")
cell.number.Mut14 <- c(cell.number.Mut14, sum(Idents(Mut14) == i))}

cat("\n\n\\pagebreak\n")
DimPlot(Mut15, reduction = "umap", label = T,label.color = "black",repel = T,label.size = 4,
        cols = ALL.colors) + 
  labs(title = "Clusters in Mut15", subtitle = 'merge all WT + Mut eight samples')

cell.number.Mut15 <- c()
cat("  \n")
for (i in levels(Idents(Mut15))){
cat(paste0('The cell number of cluster', i, ' in Mut15 is ', sum(Idents(Mut15) == i)), "\n")
cat("  \n")
cell.number.Mut15 <- c(cell.number.Mut15, sum(Idents(Mut15) == i))}

cat("\n\n\\pagebreak\n")
DimPlot(Mut16, reduction = "umap", label = T,label.color = "black",repel = T,label.size = 4,
        cols = ALL.colors) + 
  labs(title = "Clusters in Mut16", subtitle = 'merge all WT + Mut eight samples')

cell.number.Mut16 <- c()
cat("  \n")
for (i in levels(Idents(Mut16))){
cat(paste0('The cell number of cluster', i, ' in Mut16 is ', sum(Idents(Mut16) == i)), "\n")
cat("  \n")
cell.number.Mut16 <- c(cell.number.Mut16, sum(Idents(Mut16) == i))}

cat("\n\n\\pagebreak\n")

names(cell.number.WT) <- paste0(levels(Idents(WT)))
names(cell.number.Mut) <- paste0(levels(Idents(Mut)))
names(cell.number.Mut13) <- paste0(levels(Idents(Mut13)))
names(cell.number.Mut14) <- paste0(levels(Idents(Mut14)))
names(cell.number.Mut15) <- paste0(levels(Idents(Mut15)))
names(cell.number.Mut16) <- paste0(levels(Idents(Mut16)))
```


\pagebreak
# Show the top10 marker genes of each cell cluster in all 8 samples merged together
```{r message=FALSE, warning=FALSE, include=FALSE}
ALL.merge.anno.anno <- ALL.merge.anno

Idents(ALL.merge.anno.anno) <- ALL.merge.anno.anno$orig.ident
WT.anno <- subset(x = ALL.merge.anno.anno, idents = sample.name[9:12], invert=F)
Mut.anno <- subset(x = ALL.merge.anno.anno, idents = sample.name[13:16], invert=F)
Idents(ALL.merge.anno.anno) <- ALL.merge.anno.anno$seurat_clusters
Idents(WT.anno) <- WT.anno$seurat_clusters
Idents(Mut.anno) <- Mut.anno$seurat_clusters

# Show the top10 marker genes of each cell cluster.
## I can generate all the marker genes of each cluster if we need.
ALL.merge.anno.anno.markers <- FindAllMarkers(ALL.merge.anno.anno, only.pos = T)
Top10 <- ALL.merge.anno.anno.markers %>% group_by(cluster) %>% top_n(n=10, wt=avg_log2FC)

WT.anno.markers <- FindAllMarkers(WT.anno, only.pos = T)
Mut.anno.markers <- FindAllMarkers(Mut.anno, only.pos = T)
```
```{r echo=FALSE, message=FALSE, warning=FALSE}
kable(Top10) # list the top 10 marker genes in each cluster.
```

```{r message=FALSE, warning=FALSE, include=FALSE}
# Show the top100 marker genes of each cell cluster.
Top200 <- ALL.merge.anno.anno.markers %>% group_by(cluster) %>% top_n(n=200, wt=avg_log2FC)
# kable(Top100) # list the top 100 marker genes in each cluster.
write_xlsx(Top200,".\\Top200_WT-Ka.Y_markers.22.03.2024.xlsx")
write_xlsx(Top10,".\\Top10_WT-Ka.Y_markers.22.03.2024.xlsx")
```

\pagebreak
## Expression level of marker genes in WT + Mut merge.
```{r echo=FALSE, fig.height=9, fig.width=7, message=FALSE, warning=FALSE}
DoHeatmap(ALL.merge.anno.anno, features = Top10$gene, group.colors =
            ALL.colors, size = 4.5) + 
  scale_fill_gradientn(colors = heatmap.five.colors)
```

\pagebreak
## Violin plot: show example marker genes in each cluster.
```{r echo=FALSE, fig.height=2, fig.width=3, message=FALSE, warning=FALSE}
for (i in 1:8) {
Vln <- VlnPlot(ALL.merge.anno.anno, features = Top10$gene[i],pt.size=0, cols = ALL.colors)+
     theme(axis.text.x = element_text(size = 8), axis.title.y = element_text(size = 10), 
           axis.title.x = element_text(size = 0)) + theme(legend.position = "none")
print(Vln)}
```

```{r echo=TRUE}
sessionInfo()
```

